Effect of Black Hole Spin on Christodoulou Memory

### 2019-01-16 notes/Paper.tex
updated notes on paper

\documentclass[prd,preprintnumbers,twocolumn,eqsecnum,floatfix,letter]{revtex4}
\usepackage{color}
\usepackage{calc}
\usepackage{amsmath,amssymb,graphicx}
\usepackage{amssymb,amsmath}
\usepackage{tensor}
\usepackage{bm}
\usepackage{times}
\usepackage[varg]{txfonts}
\usepackage[colorlinks, pdfborder={0 0 0}]{hyperref}
\definecolor{LinkColor}{rgb}{0.75, 0, 0}
\definecolor{CiteColor}{rgb}{0, 0.5, 0.5}
\definecolor{UrlColor}{rgb}{0, 0, 0.75}
\hypersetup{linkcolor=LinkColor}
\hypersetup{citecolor=CiteColor}
\hypersetup{urlcolor=UrlColor}
\maxdeadcycles=1000
\allowdisplaybreaks
\textwidth 7.5 in 
\hoffset -1 cm 
\newcommand{\comment}[1]{\textcolor{blue}{\textit{#1}}}
\newcommand{\ashok}[1]{\textcolor{red}{\textit{Ashok:#1}}}
\newcommand{\sean}[1]{\textcolor{cyan}{\textit{Sean:#1}}}

@@

Gravitational wave (GW) memory effect leads to a permanent displacement of test masses after the gravitational wave has passed through it~\cite{Abbott_2016nmj}. The are two kinds of gravitational wave memory: linear and nonlinear. Linear memory is produced by gravitational sources that produce a net change in the time derivative of one or more of their source-multipole moments. At leading order in a Post-Newtonian (PN) expansion, the linear memory causes a net change in the GW field given by
\begin{equation}
	\Delta h^{TT}_{jk} = \frac{2}{R}\Delta\left(\ddot{\mathit{I}}^{TT}_{jk}\right) 
\end{equation} 
where $I_{ij}$ is the source mass-quadrupole moment, R is the distance source and $\Delta$ describes the net change in the quantity from early to late times. Gravitational waves with linear memory are important in case of unbound systems and have been studied studied in context of supernova explosion [add ref], asymmetric mass loss due to neutrino emission and gamma-ray-burst jets [add ref] \\

A general formula for the linear memory produced by a system of N bodies with changing mass $M_A$ or velocities $V_A$ is given by Thorn [add ref]

\begin{equation}
\Delta h^{TT}_{jk} = \Delta \sum_{A=1}^{N}\frac{4 M_A}{R\sqrt{1-v^2_A}}\left[\frac{v^j_A v^k_A}{1-\boldsymbol{v}_A\cdot \boldsymbol{N}}\right]
\end{equation}

@@

 Nonlinear memory effect also called Christodoulou memory, arises from a change in the radiative multiple moment that is sourced by the energy flux of the radiated gravitational wave. It can be understood in the following way. Consider the Einstein field equation in harmonic gauge. [add ref] 

@@

\begin{align}	

@@

is the gravitational field tensor, $g$ is the determinant of the metric $g_{\alpha\beta}$, $T^{\alpha\beta}$ is the matter stress-energy tensor, $t^{\alpha\beta}_{LL}$ is the Landau-Lifshitz pseudotensor, $\Box \equiv  -\partial^{2}_{t} + \nabla^{2}$ is the flat-space wave operator, a comma denotes a partial derivative $(,_{\mu}\equiv \partial_{\mu})$, and $\nabla^{2}$ is a flat-space Laplace operator. The Landau-Lifshitz term is the Gravitational wave stress tensor[add ref].

@@

where $\frac{dE^{gw}}{dtd\Omega}$ is the GW energy flux, $n_{j}$ is a unit radial vector, the angle-bracket mean to average over several wavelengths, $h^{TT}_{ab,j}=\bar{h}^{TT}_{ab,j}$. When we apply standard Green's function to right hand side of equation (1.1a), we obtain the following correction term to GW field [add ref]

@@

where $T_{R}$ is the retarded time. This equation shows that part of the distant GW is sourced by loss of GW energy.\\
The Christodoulou memory is an interesting consequence of the nonlinearity of general relativity. It has a clear physical interpretation : It arises from the loss of GW energy from the system and the effect of this loss on the system's radiative mass multipole moments. The nonlinear memory effect arises from nonlinear interactions at 2.5 PN order but affects the gravitational waveform at leading (0PN) order. This memory effect causes a non oscillatory shift in the amplitude of $+$ polarization, which starts at early times and slowly grows during the inspiral. The PN waveform allow us to model the slow growth of memory during the inspiral phase of coalescence but, most contribution to this nonlinear memory come from the merger phase where the PN model do not work. In this case the memory must be extracted from Numerical relativity simulation. However, it is quite challenging to extract the nonlinear memory accurately from numerical simulations of binary black holes.  Numerical relativity simulations can most accurately compute the $(l, m) = (2, 2)$ mode of the waveform; but the nonlinear memory is present in only $m=0$ mode. These $m=0$ modes tend to be much smaller in magnitude and depends more sensitively on	initial conditions than other modes. For detailed discussion on this see ref [add ref] \\
The NR simulations mostly easily compute  the curvature scalar $\Psi_4$ and decompose its values at large $R$ into a sum over spin-weighted spherical harmonic modes:

@@

In terms of the $\psi_{lm}$ modes, the leading-order Christodoulou-memory modes are smaller than the $\psi_{22}$ modes by $five$ PN orders.The relative mode magnitudes are $\psi_{22} \approx 690 \psi_{20}\approx 4.5 \times 10^4 \psi_{40}$ at $x\approx1/5$. There are other difficulties in calculating memory from numerical relativity data.Construction of the $h_{lm}$ from $\psi_{lm}$ requires two integration constants to be specified. Choosing these integration constant to be zero [add ref] causes and artificial memory and a slope in the $h_{lm}$ modes. This artificial memory is unphysical and arises from the finite size of the initial separation and extraction radius.The initial burst of junk radiation also contribute to artificial memory.\\
Numerical relativity simulations starts at a finite radius, so although it contains a most part of gravitational wave memory, PN results should be used for early inspiral phase. The significance of memory accumulated during the inspiral phase can be seen by looking at the largest memory mode $h_{20}$. Consider the leading-(Newtonian)-order piece of the largest memory mode $h_{20}$:

@@

where $U^{(mem)(1)}_{20}$ is given by the leading order piece of Eq. (3.13a) in ref [add ref], and the $T_0$ is the starting time of the simulation. For a realistic, quasicircular binary that inspirals from a large initial separation, $T_0 \rightarrow -\infty$. The NR simulations of binaries that starts at a finite initial separation would underestimate the size of the memory. For a simulation that starts with an initial separation of $10M$, the error in the memory at a harmonic coordinate separation of $5M$ is $50\%$.

\par The corrections to the leading-oder formula for the nonlinear memory's contribution to 3PN order for non-spinning binaries has been computed by by Favata []. However the effect of binary spins have not been explored these results.(\ashok{need to add why looking at spins is important})

In this paper we calculate the nonlinear memory contribution to the + waveform polarization and look at the effect of spin contribution.In Sec. II, we describe the calculation required to compute memory contribution to the post-Newtonian waveform of quasi-circular, inspiralling binaries. In Sec. III, we describe the method used extract memory contribution from NR simulations. Sec. IV we give the results which looks at the effect of spin on total nonlinear memory. We use PN expressions for memory for early time and for late time we use memory extracted from NR simulations.

@@

The mode decomposition of gravitational wave polarization is given by
\begin{equation}
	h_+ - \mathit{i}h_{\times} = \sum_{l=2}^{\infty}\sum_{m=-l}^{m=l}h^{lm} \,  _{-2}Y^{lm}(\Theta, \Phi)
\end{equation}
where $_{-2}Y^{lm}$ are spin-weighted spherical harmonics, the angles $(\Theta, \Phi)$ indicates the direction from the source to the observer. In a multipolar expansion of the GW field, the modes $h^{lm}$ are related to the radiative mass $U^{lm}$ and the current $V^{lm}$ multipole via
\begin{equation}
	h^{lm}= \frac{G}{\sqrt{2}\, R }\left[U^{lm}(T_R)-\frac{\mathit{i}}{c}V^{lm}(T_R)\right]
\end{equation}
Here $R$ is the distance from source to observer, $T_R$ is retarded time. We assume $c\,=\,G\,=\,1$ for all out calculations.
 The spin weighted spherical harmonics are defined in terms of the Wigner $d$ function by
\begin{equation}
	_{-2}Y^{lm}(\Theta, \Phi) = (-1)^2\sqrt{\frac{2l + 1}{4\pi}}d^{l}_{ms}(\Theta)e^{\mathit{i}m\Phi}
\end{equation} 
Here
\begin{align}
	d^{l}_{ms}&=\sqrt{(l+m)!(l-m)!(l+s)!(l-s)!}\\
	&\times \sum_{k=k_i}^{k_f}\frac{(-1)^k(\sin\frac{\Theta}{2})^{2k+s-1}(\cos\frac{\Theta}{2})^{1l+m-s-2k}}{k!(l+m-k)!(l-s-k)!(s-m+k)!}
\end{align}
where $k_i$ = max$(0, m-s)$ and $k_f$ = min$(l+m, l-s)$. \\
When the GW field is decomposed into modes as in [ref eqn], the nonlinear memory can be shown to yield a correction to the radiative mass multipole moments that enters at 2.5PN and higher orders.
\begin{equation}
	U^{(mem)}_{lm}=32\pi\frac{(l-2)!}{2(l+2)!}\int_{-\infty}^{T_R}dt\int d\Omega\frac{dE_{gw}}{dtd\Omega}(\Omega)Y^*_{lm}(\Omega)
\end{equation}
The radiative current moments $V_{lm}$ do not contribute to nonlinear memory. The GW energy flux can be computed from the GW stress-energy tensor and is given by [add ref]
\begin{equation}
	\frac{dE_{gw}}{dtd\Omega} = R^2 T^{gw}_{00} = \frac{R^2}{32\pi}\langle\dot{h}^{TT}_{jk}\dot{h}^{TT}_{jk}\rangle =\frac{R^2}{16\pi}\langle\dot{h}^2_{+}+ \dot{h}^2_{\times}\rangle
\end{equation}
where the angled bracket mean to average over several wavelengths. The energy flux and be written in terms of $h_{lm}$ modes using equation 2.1 in the following way:
\begin{widetext}
\begin{equation}
		\frac{dE_{gw}}{dtd\Omega}=\frac{R^2}{16\pi}\sum_{l'=2}^{\infty}\sum_{l''=2}^{\infty}\sum_{m'=-l'}^{l'}\sum_{m''=-l''}^{l''}\langle\dot{h}_{l'm'}\dot{h}^*_{l''m''}\rangle _{-2}Y^{l'm'}(\theta,\phi)_{-2}Y^{l''m''*}(\theta, \phi)
\end{equation}
\end{widetext}

The memory contribution to mass multipole moment can now be evaluated using equation 2.6 and 2.8. The angular part of the integral has the following form

\begin{equation}
	\int d\Omega _{-2}Y_{l'm'}(\theta,\phi)_{-2}Y_{l''m''}^{*}(\theta, \phi)Y_{lm}^{*}(\theta, \phi)
\end{equation}

The above integral can be evaluated as done in ref [add reference]. The result is show in appendix. \\
The time derivative of the memory mass-multipole moment is defines as [add ref] 
 $U^{(mem)(1)}_{lm}\equiv\frac{dU^{(mem)}_{lm}}{dT_R}$, which when combined using equation 2.6 and 2.8, can be written as

\begin{align}
	U_{lm}^{(mem)(1)}=&R^{2}\sqrt{\frac{2(l-2)!}{(l+2)!}}\sum_{l'=2}^{\infty}\sum_{l''=2}^{\infty}\sum_{m'=-l'}^{l'}\sum_{m''=-l''}^{l''}(-1)^{m+m'}\\
	&\times\Bigg\langle \dot{h}_{l'm'}\dot{h}_{l'm'}^{*}\Bigg\rangle G^{2-20}_{l'm'lm'-m''-m}
\end{align}

where $G^{2-20}_{l'm'lm'-m''-m}$ is the angular integral of equation 2.9 and is given in Appendix [add]. The above equation can be used to compute the first time derivative of mass-multipole moment by substituting the GW modes $h_{lm}$. When evaluating equation 2.10, we are only interested in $m=0$ modes. The $m\neq=0$ terms yield oscillatory contribution to the waveform polarizations that enter at higher PN orders then the nonoscillatory, $m=0$ terms. The non-vanishing, non-oscillatory modes are $U^{(mem)}_{l0}$ with l-even. These $U^{(mem)}_{l0}$ consist of time-integral of polynomials in $x\equiv(M\omega)^{2/3}$ (where $\omega$ is the orbital angular frequency), with each terms of the form
\begin{equation}
\int_{-\infty}^{T_{R}} x^{n}dt=\int_{-\infty}^{T_{R}}\frac{x^{n}}{\dot{x}}dx
\end{equation}
After performing the change of variable indicated in above equation 2.12, evaluation of the resulting integrals requires a model for the frequency evolution of the binary. The adiabatic evolution of the frequency (or $x$) is easily derived from energy balance ($\mathcal{L}_{GW}=-\dot{E}$). and the relation $\dot{x}=-\mathcal{L}_{GW}/(dE/dx)$. The 3.5PN orbital energy is given by equation C in ref[add aruns paper ref]
\begin{widetext}
	\begin{align}\label{eq:E}
	E&=-\frac{\eta \, M \, x}{2}\left\{1 + x \left[-\frac{3}{4}-\frac{\eta}{12}\right] + x^{3/2}\left[\left(\frac{8}{3} - \frac{4 \, \eta}{3}\right)\chisdL + \frac{8}{3}\delta\chiadL\right]\right\} \nonumber \\
	\end{align}

and the 3PN GW luminosity is given by Eq C7 to C12 in [ref]

@@

Substituting the above two equation into $\dot{x}=-\mathcal{L}_{GW}/(dE/dx)$, we get the following result

\begin{align}\label{eq:FbydE}
\frac{dx}{dt} & = \frac{64\, x^5  \, \eta}{5} \left\{1 + x \left[-\frac{743}{336}-\frac{11 \eta }{4} \right] 
+ x^{3/2} \left[4 \pi-\frac{113}{12}\delta \, \chiadL- \left(-\frac{113}{19} + \frac{19 \, \eta}{3}  \right)\chisdL \right] \right. \nonumber \\ 
\end{align}
\end{widetext}
The above equation can be used to evaluate the integral in Eq. 2.12

>>>>

\documentclass[prd,preprintnumbers,twocolumn,eqsecnum,floatfix,letter]{revtex4}
\usepackage{color}
\usepackage{calc}
\usepackage{amsmath,amssymb,graphicx}
\usepackage{amssymb,amsmath}
\usepackage{tensor}
\usepackage{bm}
\usepackage{times}
\usepackage[varg]{txfonts}
\usepackage[colorlinks, pdfborder={0 0 0}]{hyperref}
\definecolor{LinkColor}{rgb}{0.75, 0, 0}
\definecolor{CiteColor}{rgb}{0, 0.5, 0.5}
\definecolor{UrlColor}{rgb}{0, 0, 0.75}
\hypersetup{linkcolor=LinkColor}
\hypersetup{citecolor=CiteColor}
\hypersetup{urlcolor=UrlColor}
\maxdeadcycles=1000
\allowdisplaybreaks
\textwidth 7.5 in 
\hoffset -1 cm 
\newcommand{\RomanNumeralCaps}[1]{\MakeUppercase{\romannumeral #1}}
\newcommand{\comment}[1]{\textcolor{blue}{\textit{#1}}}
\newcommand{\ashok}[1]{\textcolor{red}{\textit{Ashok:#1}}}
\newcommand{\sean}[1]{\textcolor{cyan}{\textit{Sean:#1}}}

@@

Gravitational wave memory effect results in a permanent displacement of test masses after the gravitational wave has passed through an "ideal" GW detector. There are two kinds of gravitational wave memory effects: linear and nonlinear. The linear memory is produced by gravitational sources that produce a net change in the time derivative of one or more of their source-multipole moments. At leading order in a Post-Newtonian (PN) expansion, the linear memory causes a net change in the GW field given by
\begin{equation}
\Delta h^{TT}_{jk} = \frac{2}{R}\Delta\left(\ddot{\mathit{I}}^{TT}_{jk}\right) 
\end{equation} 
where $I_{ij}$ is the source mass-quadrupole moment, R is the distance to the source and $\Delta$ describes the net change in the quantity from early to late times. Gravitational waves with linear memory are important in case of unbound systems and have been studied studied in context of supernova explosion [add ref], asymmetric mass loss due to neutrino emission and gamma-ray-burst jets, \cite{Davies_Melvyn2002,Buonanno_Sigl2015,Kotake_Kei2009}
\par The linear memory produced by a system of N bodies with changing mass $M_A$ or velocities $V_A$ is given by the following general formula for by Thorn \cite{Thorne_Braginskii1987}

\begin{equation}\label{eq:1.2}
\Delta h^{TT}_{jk} = \Delta \sum_{A=1}^{N}\frac{4 M_A}{R\sqrt{1-v^2_A}}\left[\frac{v^j_A v^k_A}{1-\boldsymbol{v}_A\cdot \boldsymbol{N}}\right]
\end{equation}

@@

The other kind of memory is the nonlinear memory, which is also called Christodoulou memory. This was independently discovered by Christodoulou \cite{Christodoulou1991}, Payne \cite{Payne1894}, Blanchet and Damour \cite{Blanchet_Damour1992}. Christodoulou arises from a change in the radiative multiple moment that is sourced by the energy flux of the radiated gravitational wave. It can be understood in the following way. Consider the Einstein field equation in harmonic gauge. \cite{Thorne1980} 

@@

\begin{align}\label{eq:3.1a}	

@@

is the gravitational field tensor, $g$ is the determinant of the metric $g_{\alpha\beta}$, $T^{\alpha\beta}$ is the matter stress-energy tensor, $t^{\alpha\beta}_{LL}$ is the Landau-Lifshitz pseudotensor, $\Box \equiv  -\partial^{2}_{t} + \nabla^{2}$ is the flat-space wave operator, a comma denotes a partial derivative $(,_{\mu}\equiv \partial_{\mu})$, and $\nabla^{2}$ is a flat-space Laplace operator. The Landau-Lifshitz term is the Gravitational wave stress tensor~\cite{Isaacson1968}.

@@

where $\frac{dE^{gw}}{dtd\Omega}$ is the GW energy flux, $n_{j}$ is a unit radial vector, the angle-bracket mean to average over several wavelengths, $h^{TT}_{ab}=\bar{h}^{TT}_{ab}$. When we apply standard Green's function to right hand side of equation \ref{eq:3.1a}, we obtain the following correction term to GW field \cite{Wiseman1991}
\begin{equation}\label{eq:1.6}

@@

where $T_{R}$ is the retarded time. This equation shows that part of the distant GW is sourced by loss of GW energy. The nonlinear memory occurs in any system that radiates GWs. These systems are usually considered to have a vanishing linear memory. The nonlinear memory [Eq.\ref{eq:1.6}] can be described by equation \ref{eq:1.2}, the formula for the linear memory if the unbound object in the system are taken to be the individual gravitons with energy $E_A = M_A/(1-v_A^2)^{1/2}$ and velocities $v_A^j = c\,{n'}_A^{j}$.\cite{Thorne1992} 
\par The Christodoulou memory is an interesting consequence of the nonlinearity of general relativity, it arises from the loss of GW energy from the system and the effect of this loss on the system's radiative mass multipole moments. The nonlinear memory effect arises from nonlinear interactions at 2.5 PN order but affects the gravitational waveform at leading (0PN) order. This memory effect causes a non oscillatory shift in the amplitude of $+$ polarization, which starts at early times and slowly grows during the inspiral. The PN waveform allow us to model the slow growth of memory during the inspiral phase of coalescence, but the most contribution to this nonlinear memory come from the merger phase where the PN model do not work. In this case the memory must be extracted from Numerical relativity simulations. However, it is quite challenging to extract the nonlinear memory accurately from numerical simulations of binary black holes. Numerical relativity simulations can most accurately compute the $(l, m) = (2, 2)$ mode of the waveform; but the nonlinear memory is present in only $m=0$ mode. These $m=0$ modes tend to be much smaller in magnitude and depends more sensitively on	initial conditions than other modes. For detailed discussion on this see section $\RomanNumeralCaps{5}$ in reference \cite{Favata2009}
\par The NR simulations mostly easily compute  the curvature scalar $\Psi_4$ and decompose its values at large $R$ into a sum over spin-weighted spherical harmonic modes:

@@

In terms of the $\psi_{lm}$ modes, the leading-order Christodoulou-memory modes are smaller than the $\psi_{22}$ modes by $five$ PN orders.The relative mode magnitudes(\cite{Favata2009}) are $\psi_{22} \approx 690 \psi_{20}\approx 4.5 \times 10^4 \psi_{40}$ at $x\approx1/5$. There are many other difficulties too in calculating memory from numerical relativity data. Construction of the $h_{lm}$ from $\psi_{lm}$ requires two integration constants to be specified. Choosing these integration constant to be zero \cite{Reisswig_Pollney2011} causes an artificial memory and a slope in the $h_{lm}$ modes. This artificial memory is unphysical and arises from the finite size of the initial separation and extraction radius.The initial burst of junk radiation also contribute to artificial memory.
\par Numerical relativity simulations starts at a finite radius, so although it contains most part of gravitational wave memory, PN results should be used for early inspiral phase. The significance of memory accumulated during the inspiral phase can be seen by looking at the largest memory mode $h_{20}$. Consider the leading-(Newtonian)-order piece of the largest memory mode $h_{20}$:

@@

where $U^{(mem)(1)}_{20}$ is given by the leading order piece of Eq. (3.13a) in ref \cite{Favata2009}, and the $T_0$ is the starting time of the simulation. For a realistic, quasicircular binary that inspirals from a large initial separation, $T_0 \rightarrow -\infty$. The NR simulations of binaries that starts at a finite initial separation would underestimate the size of the memory. For a simulation that starts with an initial separation of $10M$, the error in the memory at a harmonic coordinate separation of $5M$ is $50\%$.\cite{Favata2009}

\par The corrections to the leading-oder formula for the nonlinear memory's contribution to 3PN order for non-spinning binaries has been computed by by Favata \cite{Favata2009}. However the effect of binary spins have not been explored these results.(\ashok{need to add why looking at spins is important})   
\par In this paper we calculate the nonlinear memory contribution to the + waveform polarization and look at the effect of spin contribution. In Sec. II, we describe the calculation required to compute memory contribution to the post-Newtonian waveform of quasi-circular, inspiralling binaries. In Sec. III, we describe the method used to extract memory contribution from NR simulations. Sec. IV we give the results which looks at the effect of spin on total nonlinear memory. We use PN expressions for memory for early time and for late time we use memory extracted from NR simulations.

@@

The mode decomposition of gravitational wave polarization is given by
\begin{equation}\label{eq:2.1}
	h_+ - \mathit{i}h_{\times} = \sum_{l=2}^{\infty}\sum_{m=-l}^{m=l}h^{lm} \,  _{-2}Y^{lm}(\Theta, \Phi)
\end{equation}
where $_{-2}Y^{lm}$ are spin-weighted spherical harmonics, the angles $(\Theta, \Phi)$ indicates the direction from the source to the observer. In a multipolar expansion of the GW field, the modes $h^{lm}$ are related to the radiative mass $U^{lm}$ and the current $V^{lm}$ multipole via
\begin{equation}\label{eq:2.2}
	h^{lm}= \frac{G}{\sqrt{2}\, R }\left[U^{lm}(T_R)-\frac{\mathit{i}}{c}V^{lm}(T_R)\right]
\end{equation}
Here $R$ is the distance from source to observer, $T_R$ is retarded time. We assume $c\,=\,G\,=\,1$ for all out calculations.
 The spin weighted spherical harmonics are defined in terms of the Wigner $d$ function by
\begin{equation}\label{eq:2.3}
	_{-2}Y^{lm}(\Theta, \Phi) = (-1)^2\sqrt{\frac{2l + 1}{4\pi}}d^{l}_{ms}(\Theta)e^{\mathit{i}m\Phi}
\end{equation} 
Here
\begin{align}\label{eq:2.4}
	d^{l}_{ms}&=\sqrt{(l+m)!(l-m)!(l+s)!(l-s)!}\\
	&\times \sum_{k=k_i}^{k_f}\frac{(-1)^k(\sin\frac{\Theta}{2})^{2k+s-1}(\cos\frac{\Theta}{2})^{1l+m-s-2k}}{k!(l+m-k)!(l-s-k)!(s-m+k)!}
\end{align}
where $k_i$ = max$(0, m-s)$ and $k_f$ = min$(l+m, l-s)$. \\
When the GW field is decomposed into spin-weighted spherical harmonic modes, the nonlinear memory can be shown to yield a correction to the radiative mass multipole moments that enters at 2.5PN and higher orders. For more detail about GW multipole expansion \cite{Thorne1980}. The correction to the radiative mass multipole moments is given by
\begin{equation}\label{eq:2.6}
	U^{(mem)}_{lm}=32\pi\frac{(l-2)!}{2(l+2)!}\int_{-\infty}^{T_R}dt\int d\Omega\frac{dE_{gw}}{dtd\Omega}(\Omega)Y^*_{lm}(\Omega)
\end{equation}
The radiative current moments $V_{lm}$ do not contribute to nonlinear memory. The GW energy flux can be computed from the GW stress-energy tensor and is given by \cite{Favata2010}
\begin{equation}\label{eq:2.7}
	\frac{dE_{gw}}{dtd\Omega} = R^2 T^{gw}_{00} = \frac{R^2}{32\pi}\langle\dot{h}^{TT}_{jk}\dot{h}^{TT}_{jk}\rangle =\frac{R^2}{16\pi}\langle\dot{h}^2_{+}+ \dot{h}^2_{\times}\rangle
\end{equation}
where the angled bracket mean to average over several wavelengths. The energy flux can be written in terms of $h_{lm}$ modes using equation \ref{eq:2.1} in the following way:
\begin{widetext}
\begin{equation}\label{eq:2.8}
		\frac{dE_{gw}}{dtd\Omega}=\frac{R^2}{16\pi}\sum_{l'=2}^{\infty}\sum_{l''=2}^{\infty}\sum_{m'=-l'}^{l'}\sum_{m''=-l''}^{l''}\langle\dot{h}_{l'm'}\dot{h}^*_{l''m''}\rangle _{-2}Y^{l'm'}(\theta,\phi)_{-2}Y^{l''m''*}(\theta, \phi)
\end{equation}
\end{widetext}

The memory contribution to the mass multipole moment can now be evaluated using equation \ref{eq:2.6} and \ref{eq:2.8}. The angular part of the integral has the following form

\begin{equation}\label{eq:2.9}
	\int d\Omega _{-2}Y_{l'm'}(\theta,\phi)_{-2}Y_{l''m''}^{*}(\theta, \phi)Y_{lm}^{*}(\theta, \phi)
\end{equation}

The above integral can be evaluated as done in section III of  Favata\cite{Favata2009}. The results are also shown in appendix ~\ref{Appendix A}. The time derivative of the memory mass-multipole moment is defines as $U^{(mem)(1)}_{lm}\equiv\frac{dU^{(mem)}_{lm}}{dT_R}$ \cite{Favata2009}, which when combined using equation \ref{eq:2.6} and \ref{eq:2.8} and Appendix \ref{Appendix A}, can be written as

\begin{align}\label{eq:2.10}
	U_{lm}^{(mem)(1)}=&R^{2}\sqrt{\frac{2(l-2)!}{(l+2)!}}\sum_{l'=2}^{\infty}\sum_{l''=2}^{\infty}\sum_{m'=-l'}^{l'}\sum_{m''=-l''}^{l''}(-1)^{m+m'}\nonumber\\
	&\times\Bigg\langle \dot{h}_{l'm'}\dot{h}_{l'm'}^{*}\Bigg\rangle G^{2-20}_{l'm'lm'-m''-m}
\end{align}

where $G^{2-20}_{l'm'lm'-m''-m}$ is the angular integral of equation \ref{eq:2.9} and is given in Appendix \ref{Appendix A}. The above equation can be used to compute the first time derivative of mass-multipole moment by substituting the GW modes $h_{lm}$. When evaluating equation \ref{eq:2.10}, we are only interested in $m=0$ modes. The $m\neq0$ terms yield oscillatory contribution to the waveform polarizations that enter at higher PN orders then the nonoscillatory, $m=0$ terms. The non-vanishing, non-oscillatory modes are $U^{(mem)}_{l0}$ with l-even. These $U^{(mem)}_{l0}$ consist of time-integral of polynomials in $x\equiv(M\omega)^{2/3}$ (where $\omega$ is the orbital angular frequency), with each terms of the form
\begin{equation}\label{eq:2.11}
\int_{-\infty}^{T_{R}} x^{n}dt=\int_{-\infty}^{T_{R}}\frac{x^{n}}{\dot{x}}dx
\end{equation}
After performing the change of variable indicated in above equation \ref{eq:2.11}, evaluation of the resulting integrals requires a model for the frequency evolution of the binary. The adiabatic evolution of the frequency (or $x$) is easily derived from energy balance ($\mathcal{L}_{GW}=-\dot{E}$). and the relation $\dot{x}=-\mathcal{L}_{GW}/(dE/dx)$. The 3.5PN orbital energy and flux are given in Appendix C of reference \cite{Arun2009}. Here we who only upto 1.5 PN terms, the complete expression are show in Appendix \ref{Appendix:C1}. The 1.5 PN expression for energy is given by 
\begin{widetext}
	\begin{align}\label{eq:E}
	E&=-\frac{\eta \, M \, x}{2}\left\{1 + x \left[-\frac{3}{4}-\frac{\eta}{12}\right] + x^{3/2}\left[\left(\frac{8}{3} - \frac{4 \, \eta}{3}\right)\chisdL + \frac{8}{3}\delta\chiadL\right]\right\} \nonumber \\
	\end{align}

and the 1.5 PN GW luminosity is given by

@@

Substituting the above two equation into $\dot{x}=-\mathcal{L}_{GW}/(dE/dx)$, and expanding upto 1.5 PN gives the following expression.

\begin{align}\label{eq:2.14}
\frac{dx}{dt} & = \frac{64\, x^5  \, \eta}{5} \left\{1 + x \left[-\frac{743}{336}-\frac{11 \eta }{4} \right] 
+ x^{3/2} \left[4 \pi-\frac{113}{12}\delta \, \chiadL- \left(-\frac{113}{19} + \frac{19 \, \eta}{3}  \right)\chisdL \right] \right. \nonumber \\ 
\end{align}
\end{widetext}
The results upto 3.5 PN are shows in Appendix \ref{Appendix:C}.
The equation \ref{eq:2.14} can be used change the integration with respect to time into integration with respect to $x$ as shown in equation in \ref{eq:2.11}. The expressions for $h_{lm}$ whose first time derivatives appear equation \ref{eq:2.10} are directly taken from reference \cite{Arun2009}. The PN expression for $h_{lm}$ there are computed for precessing binaries(Equations 4.17 in ref \cite{Arun2009}), but here we are only interested in aligned spin case. The structure of modes for non-precessing binaries can be obtained by aligning the orbital angular momentum with the total angular momentum($\iota = 0, \alpha = \pi$). These results for a special case of non-precessing binaries is shown in Appendix \ref{Appendix:D}. We show the $h_{lm}$ expressions for positive $m$ values, but the results for negative $m$ values can be obtained by the relation, $h_{l, -m} = (-1)^l \bar{h}_{lm}$. This results is not in general true for precessing binaries, but it continues to hold true of non-pressing case.

\begin{widetext}
\subsection{RESULTS: MEMORY CONTRIBUTION TO THE POST-NEWTONIAN WAVEFORM OF QUASI-CIRCULAR, INSPIRALLING BINARIES}
The memory contribution to the spin-weighted spherical-harmonic modes of the polarization waveform [Eq. \ref{eq:2.2}]. These quantities are related via 
\begin{equation}
	h_{l0}^{(mem)} = \frac{\alpha}{\sqrt{2} \, R} U_{l0}^{(mem)} = 8 \sqrt{\frac{\pi}{5}}\frac{\eta M x}{R} 
\end{equation}
where we have followed the notation of Sec. 9 of Ref. \cite{Blanchet_Iyer2008}. The notational parameter $\alpha$ accounts for the two commonly used choices for the polarization triad 
\begin{equation}
\alpha = \left\{
\begin{array}{ll}
(+1) & \text{for the Kidder  \cite{Kidder2008} convention} \\
(-1) & \text{for the Blanchet et al. \cite{Blanchet_Iyer2008}   convention}\\
\end{array} 
\right.
\end{equation}

The results of polarization modes in terms of $\hat{H}_{l0}$ are:
\begin{subequations}
\begin{align}\label{eq:H20}
\hat{H}_{20} &= \alpha \frac{5}{14\sqrt{6}}\left\{ 1 + x\left[-\frac{4075}{4032} + \frac{67 \, \eta}{48}\right]+ x^{3/2}\left[\left(\frac{7}{480} +\frac{27 \, \delta}{10} - \frac{7 \,\eta}{120}\right)\chiadL+\left(\frac{27}{10} +\frac{7 \, \delta^3}{480} - \frac{2 \, \eta}{15}\right)\chisdL\right]\right. \nonumber \\
& \quad \left. x^2\left[-\frac{151877213}{67060224} -\frac{123815 \, \eta}{44352} +\frac{205 \, \eta^2}{352} +\left(-\frac{163}{96} +\frac{161 \, \eta}{24}\right)\chiaSqr + \left(-\frac{27 \, \delta}{8} -\frac{\delta^3}{48}\right)\chisDchia + \left(-\frac{27}{16}-\frac{\delta^4}{96} -\frac{\eta}{12}\right)\chisSqr\right]\right. \nonumber \\
& \quad \left. x^{5/2}\left[-\frac{253 \, \pi}{336}+ \frac{253 \, \pi \, \eta}{84} +\left\{\frac{631}{32256} + \frac{788471\,\delta}{84672} +\left(-\frac{5}{126}-\frac{991\,\delta}{1008}\right)\eta-\frac{311\,\eta^2}{2016}\right\}\chiadL\right.\right. \nonumber\\
& \qquad\left. +\left\{\frac{788471}{84672}-\frac{631\, \delta^3}{32256} -\left(-\frac{36509}{2646} + \frac{311\,\delta^3}{8064}\right)\eta+\frac{853\,\eta^2}{252}\right\}\chisdL\right] \nonumber \\
& \quad  x^3\left[-\frac{4397711103307}{532580106240} + \left( \frac{700464542023}{13948526592} - \frac{205 \, \pi^2}{96}\right)\eta + \frac{69527951 \, \eta^2}{166053888} + \frac{1321981 \, \eta^3}{5930496} -  \left(-\frac{7\,\pi}{256} + \frac{275\,\pi\,\delta}{12} \right.\right. \nonumber \\
& \qquad \left. \frac{7 \, \pi\,\delta}{256}\right)\chiadL + \left\{-\frac{42787}{32256} + \frac{791\,\delta}{9216} +\frac{113\,\delta^2}{576} + \left(\frac{354035}{32256} -\frac{791\,\delta}{2304}\right)\eta - \frac{2533\,\eta^2}{128}\right\}\chiaSqr\nonumber\\
&\qquad \left.\left. \left\{\frac{275\,\pi}{12} -\frac{7\,\pi\,\delta^3}{256}-\frac{56\,\pi\eta}{3} + \left(\frac{791}{9216} -\frac{47869\,\delta}{21504}-\frac{743\,\delta^3}{21504} + \frac{791\,\delta^4}{9216} + \left(-\frac{77}{192}+\frac{87269\, \delta}{2304}-\frac{11\, \delta^3}{256}\right)\eta + \frac{133\,\eta^2}{576}\right)\chiadL\right\}\chisdL\right.\right.\nonumber\\
& \qquad \left.\left.\left\{-\frac{47869}{43008} + \frac{791\,\delta^3}{9216} -\frac{743\,\delta^4}{43008} + \left(\frac{57677}{1792} -\frac{133\,\delta^3}{2304} -\frac{11\,\delta^4}{512}\right)\eta - \frac{15029\,\eta^2}{1152}\right\}\chisSqr
\right]\right\}
\end{align}
